Epistemic Semantics in Guarded String Models

### 2020-06-23 tex/draft.tex
Edits at start of Section 3

@@

With modalities and propositional negation added, the signature of $n$-agent
epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
\ex{interpretation1} illustrates the application of the language with some natural
language sentences, and their representations in the logical language.  While there
is no constant for the set of worlds, the term $\neg 0$ (the complement of the empty
set) can be used as one.  Where $h$ is the primitive test ``it' heads'' from the coin
example, $(\neg 0)h$ designates the set of worlds where it is heads.  The reason is
that $h$ corresponds to the set of atoms that evaluate $h$ to true. Then the product
term $(\neg 0)h$ restricts the set of worlds to the set of guarded strings that end
with and atom that evaluate $h$ to True.

\begin{example}   \label{interpretation1}
\begin{tabular}[t]{ll}
  It's heads. & It's tails. \\
  $(\neg 0)h$ & $(\neg 0)t$ \\ \hline
\end{tabular}
\end{example}  
\vspace{-3mm}
\hspace{2mm}\begin{tabular}{l}
Amy doesn't know that it's heads.  \\
$\neg {\Box}_a (\neg 0)h$\\ \hline
Amy knows whether it's tails. \\
$\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t$ \\ \hline
It's heads and Bob knows that it's tails. \\
$(\neg 0)h + \Box_a(\neg 0)\neg t$ \\  \hline
Bob knows that Amy knows whether it's tails. \\
$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$ \\ \hline
\end{tabular}
%% (\neg 0)t

@@

  {\em Consider adding the syntax of tests including $\bar{}$. Consider making product implicit,
    in place of $\mathbin{;}$.}

  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
  \multicolumn 2 l {\textrm{terms}} & e \in \mathsf{P} \\
  p,q & ::=& e~|~0~|~1~|~\phi~|~p+p~|~p;p~|~p^*~|\neg p~|\Diamond_a~p
  \end{array}\]

@@

We interpret each term in as a set of guarded strings as indicated in
Figure~\ref{fig:interp}. The function $\sem{ p
}^{\varphi,{\mathcal{E}}}$ interprets a term $p$ as a set of
guarded strings consistent with state constraints $\varphi$ and effect
relations ${\mathcal{E}^\varphi}$. A bare event $e$ is interpreted as
the set $\hat{\mathcal E}^\varphi(e)$. The constant $0$ is interpreted
as the empty set.  The constant $1$ is interpreted as the set
$\mathcal A^\varphi_{\mathsf B}$, i.e. $\{h\bar t, \bar h t\}$ in the
coin example. The operation $p + q$ is interpreted as the union of the
interpretations of $p$ and $q$. The operation $p;q$ interpreted as the
fusion product of the interpretations of $p$ and $q$ raised to sets:
$X \diamond Y = \{ x \diamond y \mid x \in X, y \in Y, x\diamond
y~\textit{defined}\}$.  The operation $p*$ is Kleene star, which is
interpreted to be least upper bound of the iterated fusion product of
the denotation of $p$ with itself. Subsets of $1$ indicated by
propositions $\psi$ are also elements of $K$, and these form the
Boolean algebra of tests, denoting all atoms admitted by the
constraint $\varphi$ and the boolean test $\psi$.

@@

>>>>

\newcommand{\event}{\sf{P}}
\newcommand{\test}{\sf{B}}

@@

With modalities and propositional negation added, the signature of $n$-agent
epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language, but
using juxtaposition in place of the product symbol $\cdot$.
As illustrated in \ex{interpretation1}, terms in this language are used to
represent the proposositional semantic values of English sentences.\footnote{See Sections
6 and 7 for explanation and further examples.  }

\begin{example}   \label{interpretation1}
\begin{tabular}[t]{ll}
Bob knows that Amy knows whether it's tails.  \\
$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$
\end{tabular}
\end{example}

@@

  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
  \multicolumn{3}{l}{\mbox{state terms $\sigma$ as in Figure 1}} \\
  \multicolumn 2 l {\textrm{events}} & e \in \mathsf{P} \\
  p,q & ::=& e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p
  \end{array}\]

@@

Each term in the language is interpreted as a set of guarded strings.  Where $p$ is
a term, $\sem{p}^{\test,\event,\varphi,\zeta}$ is the interpretation of the term.
The superscript captures dependence on an Epik specification.

interprets a term $p$ as a set of
guarded strings consistent with state constraints $\varphi$ and effect
relations ${\mathcal{E}^\varphi}$. A bare event $e$ is interpreted as
the set $\hat{\mathcal E}^\varphi(e)$. The constant $0$ is interpreted
as the empty set.  The constant $1$ is interpreted as the set
$\mathcal A^\varphi_{\mathsf B}$, i.e. $\{h\bar t, \bar h t\}$ in the
coin example. The operation $p + q$ is interpreted as the union of the
interpretations of $p$ and $q$. The operation $p;q$ interpreted as the
fusion product of the interpretations of $p$ and $q$ raised to sets:
$X \diamond Y = \{ x \diamond y \mid x \in X, y \in Y, x\diamond
y~\textit{defined}\}$.  The operation $p*$ is Kleene star, which is
interpreted to be least upper bound of the iterated fusion product of
the denotation of $p$ with itself. Subsets of $1$ indicated by
propositions $\psi$ are also elements of $K$, and these form the
Boolean algebra of tests, denoting all atoms admitted by the
constraint $\varphi$ and the boolean test $\psi$.

@@

\newcommand{\epikspec}{\test,\event,\varphi,\zeta}

\begin{figure}
  \[\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ e }^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
  \sem{ \psi }^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
  \sem{{p+q}}^{\varphi,{\mathcal{E}}} && \sem{{p}}^{\varphi,{\mathcal{E}}} \cup \sem{ q }^{\varphi,{\mathcal{E}}}\\
  \sem{{p;q}}^{\varphi,{\mathcal{E}}} && \sem{ p }^{\varphi,{\mathcal{E}}} \diamond \sem{{q}}^{\varphi,{\mathcal{E}}}\\
  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}

  \end{array}
  \]
  \caption{Interpretation of Epik terms as sets of guarded strings}
  \label{fig:interp2}
\end{figure}
